---
title: "POXC"
author: "Sistla, Seeta and Pold, Grace"
date: "`r Sys.Date()`"
output: html_document
--- 
#subset stock and abs separately 
#plot linear regression 
#have weights of soil (in kg)

#POXC (mg/kg soil) = [0.02 mol/L - (a+b * Abs)] * (9000 mg C/ mol)* (0.02 L solution/Wt)
# a = intcp of std curve, b = slope. Abs = unknown abs, Wt = weight of air-dried soil sample in kg

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/ssistla/Desktop')
library(lme4)
require(readxl)
require(reshape2)
require(melt)
require(ggplot2)
```

##standard regression code
```{r}
ggplotRegression <- function (fit) {

require(ggplot2)

ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  labs(title = paste("Adj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                     "Intercept =",signif(fit$coef[[1]],5 ),
                     " Slope =",signif(fit$coef[[2]], 5),
                     " P =",signif(summary(fit)$coef[2,4], 5)))
}
```

#load data files - plate plan, tecan data, melt POXC data and merge with plate plan
 
```{r}
plateplan<-read_excel("biogeochem_example.xlsx", 
                      sheet = "POXC_plateplan")[,1:4]

plateplan<-na.omit(plateplan)

POXCweight<-read_excel("biogeochem_example.xlsx", 
                 sheet = "POXC_Weights")

POXCdata<-read_excel("biogeochem_example.xlsx", 
                       sheet = "POXC_tecan")[26:33,1:13]

names(POXCdata)<-c("row", 1:12)
POXCdata[,2:13]<-lapply(POXCdata[,2:13],as.numeric)

POXCdata_melt<-melt(POXCdata)
names(POXCdata_melt)[2]<-"column"
POXCdata_melt<-merge(POXCdata_melt, plateplan, by=c("row", "column"))
```

##extract POXC standards, create std curve, remove NAs and aggregate means by sample ID
```{r}
stds<-subset(POXCdata_melt, `sample position`=="standard")
stds$Soil_ID<-as.numeric(as.character(stds$Soil_ID))
stdcurve<-summary(lm(Soil_ID~value, stds))
 
ggplotRegression(lm(Soil_ID~value, stds))
 
slope<-stdcurve$coefficients[2]
intercept<-stdcurve$coefficients[1]

POXCdata_melt0 <- na.omit(POXCdata_melt)

means<-aggregate(value~Soil_ID, POXCdata_melt0, "mean")
```
 
Calculate POXC (mg/kg soil) by  air dry  soil weight, corrected by slope, intercept, and value (absorbance)
 
```{r}

POXC_means<- merge(POXCweight, means, by.x = "Soil_ID", by.y = "Soil_ID")

POXC_means$POXC_mgkgsoil <- (0.02 - (intercept + (slope * POXC_means$value))) * (9000)* (0.02/POXC_means$`Weight (kg)`)
write.csv(POXC_means, "FinalPOXC_EXAMPLE.csv")

```
 
Plot POXC data
```{r , echo=FALSE}
POXCplot<-ggplot(POXC_means,
       mapping= aes(x=Plastic_Association,
                    y=POXC_mgkgsoil, fill=Plastic_Association))+geom_boxplot()+ theme_bw()+theme(panel.grid = element_blank(), legend.position = "none")+theme(panel.grid = element_blank()) +scale_fill_manual(values=c("hotpink3", "darkolivegreen4")) + labs(x = '', y = expression(POX~C~(frac(mg~C,kg~soil))))
POXCplot
 
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
