---
title: "Ammonium"
author: "Sistla, Seeta and Pold, Grace"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
 knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/ssistla/Desktop')
library(lme4)
require(readxl)
require(reshape2)
require(ggplot2)
require(qqplotr)
library(nlme)
library(lmerTest)
library(car)
library(ggstatsplot)
library(lubridate)
library(ggpubr)
library(tidyverse)
require(readxl)
require(reshape2)
```

##Standard curve regression chunk

```{r}
ggplotRegression <- function (fit) {

require(ggplot2)

ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  labs(title = paste("Adj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                     "Intercept =",signif(fit$coef[[1]],5 ),
                     " Slope =",signif(fit$coef[[2]], 5),
                     " P =",signif(summary(fit)$coef[2,4], 5)))
}
```


##import NH4 data from TECAN, plate plan for NH4, soil moisture
```{r }

plateplanNH4<-read_excel("biogeochem_example.xlsx", 
                         sheet = "NH4_plateplan") [,1:4]

plateplanNH4<-na.omit(plateplanNH4)

NH4<-read_excel("biogeochem_example.xlsx", 
                         sheet = "NH4_tecan")[26:33,]
names(NH4)<-c("row", 1:12)
NH4[,2:13]<-lapply(NH4[,2:13],as.numeric)

NH4_melt<-melt(NH4)
names(NH4_melt)[2]<-"column"
NH4_melt<-merge(NH4_melt, plateplanNH4, by=c("row", "column"))




Soil_Moisture <- read_excel("biogeochem_example.xlsx", 
                            sheet = "Soil_Moisture") [,c(1, 6:12)]
Soil_Moisture_ave <- Soil_Moisture %>%                                     
  group_by(Soil_ID) %>%
  summarize('Average_DryMass/WetMass' = mean(`Dry Mass/Wet Mass`)) %>% 
  as.data.frame()
Soil_Moisture_ave
```

##extract standards from NH4  plate
#check the slope and R2 visually by plotting, save skope and intercept to calculate ppm NH4
#subset samples that are not standards from NH4_melt data drame and caclulate ppm NH4 from standard curve!


```{r}
 
stds<-subset(NH4_melt, `sample position`=="standard")
stds$Soil_ID<-as.numeric(as.character(stds$Soil_ID))
stdcurve<-summary(lm(Soil_ID~value, stds))
ggplotRegression(lm(Soil_ID~value, stds))
plot(Soil_ID~value, stds)

slope<-stdcurve$coefficients[2]
intercept<-stdcurve$coefficients[1]

samples<-subset(NH4_melt, !`sample position`=="standard")
samples$ppmNH4<-samples$value*slope+intercept
```

#find means for each sample:
# To convert to µg NO 3 g -1 of soil, first, note that ppm is the same as µg ml -1 . 
#Multiply µg ml -1 NO3 from the microplate assay by the ml of extractant (KCl) solution and divide by g dry soil. Divide by 1000 to get to mg NO3 / g soil
#subtract out extract blank (neg control)
```{r}

meansNH4<-aggregate(ppmNH4~Soil_ID, samples, "mean")
meansNH4$mgNH4_perExtraction<-meansNH4$ppmNH4*20/1000
meansNH4$mgNH4_perExtraction<-meansNH4$mgNH4_perExtraction-subset(meansNH4, Soil_ID=="NEG")$mgNH4_perExtraction
```


#metadata (this merge will also get rid of empty wells and negative control wells)
#merge metadata with means, merge means with soil moisture ave
#convert to mg NH4/kg dry soil
#convert to mg NH4-N/kg dry soil (molecular weight of N/NH4 =14.0067 /18.03806 = 0.7765 
#check range is normative (0-5 mg NH4-N/kg soil)

```{r}
metadata<-read_excel("biogeochem_example.xlsx", 
                         sheet = "extractable_nuts_weight") [,1:6]
 
meansNH4<-merge(meansNH4, metadata, by.x="Soil_ID", by.y="Soil_ID")
meansNH4<-merge(meansNH4,Soil_Moisture_ave,by = "Soil_ID")

meansNH4$mg_g_wetSoil<-meansNH4$mgNH4_perExtraction/meansNH4$'Mass (g)'
meansNH4$mg_g_drySoil<-meansNH4$mg_g_wetSoil/meansNH4$'Average_DryMass/WetMass'
meansNH4$mgNH4_kg_drySoil<- meansNH4$mg_g_drySoil*1000
meansNH4$mgNH4_N_kg_drySoil<- meansNH4$mgNH4_kg_drySoil *(14.0067 /18.03806)
write.csv(meansNH4, "FinalNH4_EXAMPLE.csv")
```


#NO3 Plot by plastic association


```{r} 
NH4plot<- ggplot((meansNH4), aes(x=Plastic_Association,y=mgNH4_N_kg_drySoil))+ geom_boxplot()+
  theme_bw() + 
  theme(panel.grid = element_blank(), legend.position = "none")+
  labs(x = '', y = (expression(soil~nitrate~(mg~NH[4]~-N~~kg^{"-1"}~soil))))
NH4plot
```
 
 