---
title: "NO3"
author: "Sistla, Seeta and Pold, Grace"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
 knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/ssistla/Desktop')
library(lme4)
require(readxl)
require(reshape2)
require(ggplot2)
require(qqplotr)
library(nlme)
library(lmerTest)
library(car)
library(ggstatsplot)
library(lubridate)
library(ggpubr)
library(tidyverse)
require(readxl)
require(reshape2)
```



##import NO3 data from TECAN, plate plan for NO3, soil moisture
```{r }

plateplanNO3<-read_excel("biogeochem_example.xlsx", 
                         sheet = "NO3_plateplan") [,1:4]

plateplanNO3<-na.omit(plateplanNO3)

NO3<-read_excel("biogeochem_example.xlsx", 
                         sheet = "NO3_tecan")[26:33,]
names(NO3)<-c("row", 1:12)
NO3[,2:13]<-lapply(NO3[,2:13],as.numeric)

NO3_melt<-melt(NO3)
names(NO3_melt)[2]<-"column"
NO3_melt<-merge(NO3_melt, plateplanNO3, by=c("row", "column"))




Soil_Moisture <- read_excel("biogeochem_example.xlsx", 
                            sheet = "Soil_Moisture") [,c(1, 6:12)]
Soil_Moisture_ave <- Soil_Moisture %>%                                     
  group_by(Soil_ID) %>%
  summarize('Average_DryMass/WetMass' = mean(`Dry Mass/Wet Mass`)) %>% 
  as.data.frame()
Soil_Moisture_ave
```

##extract standards from NO3 plate
#check the slope and R2 visually by plotting, save skope and intercept to calculate ppm NO3
#subset samples that are not standards from NO3_melt data drame and caclulate ppm NO3 from standard curve!


```{r}
 
stds<-subset(NO3_melt, `sample position`=="standard")
stds$sampleID<-as.numeric(as.character(stds$Soil_ID))
stdcurve<-summary(lm(Soil_ID~value, stds))
plot(Soil_ID~value, stds)



slope<-stdcurve$coefficients[2]
intercept<-stdcurve$coefficients[1]

samples<-subset(NO3_melt, !`sample position`=="standard")
samples$ppmNO3<-samples$value*slope+intercept
```

#find means for each sample:
# To convert to µg NO 3 g -1 of soil, first, note that ppm is the same as µg ml -1 . 
#Multiply µg ml -1 NO3 from the microplate assay by the ml of extractant (KCl) solution and divide by g dry soil. Divide by 1000 to get to mg NO3 / g soil
#subtract out extract blank (neg control)
```{r}

meansNO3<-aggregate(ppmNO3~Soil_ID, samples, "mean")
meansNO3$mgNO3_perExtraction<-meansNO3$ppmNO3*20/1000
meansNO3$mgNO3_perExtraction<-meansNO3$mgNO3_perExtraction-subset(meansNO3, Soil_ID=="NEG")$mgNO3_perExtraction
```


#metadata (this merge will also get rid of empty wells and negative control wells)
#merge metadata with means, merge means with soil moisture ave
#convert to mg NO3/kg dry soil
#convert to mg NO3-N/kg dry soil (molecular weight of N/NO3 = 0.22)
#check range is normative (0-50 mg NO3-N/kg soil)

```{r}
metadata<-read_excel("biogeochem_example.xlsx", 
                         sheet = "extractable_nuts_weight") [,1:6]
 
meansNO3<-merge(meansNO3, metadata, by.x="Soil_ID", by.y="Soil_ID")
meansNO3<-merge(meansNO3,Soil_Moisture_ave,by = "Soil_ID")

meansNO3$mg_g_wetSoil<-meansNO3$mgNO3_perExtraction/meansNO3$'Mass (g)'
meansNO3$mg_g_drySoil<-meansNO3$mg_g_wetSoil/meansNO3$'Average_DryMass/WetMass'
meansNO3$mgNO3_kg_drySoil<- meansNO3$mg_g_drySoil*1000
meansNO3$mgNO3_N_kg_drySoil<- meansNO3$mgNO3_kg_drySoil *(14.0067/62.0037)
write.csv(meansNO3, "FinalNO3_EXAMPLE.csv")
```


#NO3 Plot by plastic association


```{r} 
NO3plot<- ggplot((meansNO3), aes(x=Plastic_Association,y=mgNO3_N_kg_drySoil))+ geom_boxplot()+
  theme_bw() + 
  theme(panel.grid = element_blank(), legend.position = "none")+
  labs(x = '', y = (expression(soil~nitrate~(mg~NO[3]~-N~~kg^{"-1"}~soil))))
NO3plot
```
 
 